---
title: "Introduction to R Markdown"
author: 
    name: "Jing Xu" 
    affiliation: "Applied Bioinformatics (B330), DKFZ"
    email: "jing.xu@dkfz.de"
        
date: "2019-03-08"
output: BiocStyle::html_document

---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# It's possible to change your working directory here
#knitr::opts_knit$set(root.dir = 'Your working directory') 

# load necessary packages in setup chunk
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("BiocStyle", version = "3.8")
library(BiocStyle)
```

# About R Markdown

The document format "R Markdown" was first introduced in the `knitr` package. Markdown is a simple formatting syntax for authoring HTML, PDF, beamer presentation and MS Word documents. It also allows you to work with different languages like bash and python.

*  You need to download and install R studio at <https://www.rstudio.com/products/rstudio/download/>.  
* Install R markdown package by `install.packages("rmarkdown")`, in this file, we use the "BiocStyle html_document"" for a nice looking report, therefore we need to install the `BiocStyle` package from Bioconductor, see code chunk "setup"
* Create a R markdown file from the top dropdown manual: File --> New file --> R Markdown  
* Write all the text and code in the file, You can also include multiple languages in the RMD file including R, Python, bash and SQL.  
* Create a code chunk by clicking the "Insert" button on the right corner, write your code and click "Run" or "Play" button to check the output in the current code chunk.
* When you finish your writing, click the **Knit** button, a document will be generated that includes both content as well as the output of any embedded code chunks within the document.   

R markdown allows you to document all details in the report to ensure reproducibility of your work. 

## Check out the examples provided in the RMD template
- You can embed an R code chunk like this:

```{r cars}
summary(cars)

```

- Including Plots

You can also embed plots without showing the code, for example:

```{r pressure, echo=FALSE, warning=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


# Objectives
- Learn how to prepare different types of R Markdown files for various uses
- Learn about writing and running R Markdown file from your assignment

At the end of the course, you need to complete the programming assignment in a semi-prepared RMD file and run it to generate a HTML report. It is required to hand in this file to get credits for the course.  


# More information: 
- Check R Markdown manuals from R studio website <http://rmarkdown.rstudio.com>
- Book: [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)
- R Markdown cheet sheet: check your handouts or download it from [RMD cheatsheet2.0](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf) 


# Document versions for reproducibility
You need to know which versions of R and packages you used. Document the details in the file!
```{r}
sessionInfo()
```

